### 2012-01-23 paper_draft.tex
Added new commenting system taken from here: http://stackoverflow.com/questions/539791/tips-for-collaboratively-editing-a-latex-document/539848#539848; changed already present comments

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

@@

The openSNP project (http://opensnp.org) is an open, crowdsourced online platform for both DTC customers and researchers. DTC customers are encouraged to share their genotyping results and connect them with phenotypes to enable GWAS for researchers or any other reuse in a scientific context. 
\begin{comment}
FZ: That is a good summary, but as Bastian pointed out the reuse is not only
enabled in a scientific context?
\end{comment}

\begin{comment}
FZ:
\end{comment}
\subsubsection*{Sharing genotypic information}
Currently users can upload their genotypings from the companies 23andme, deCODEme and Family Tree to/via a webinterface to the openSNP 
project. The uploaded data is published under the Creative Commons Zero-license, 
which - in accordance with the Panton Principles (doi:10.1371/journal.pbio.1001195) - 
allows a complete reuse of the data without any constraints.
\begin{comment}
FZ: If that is true, rephrase sentence above
FZ: Made two sentences out of one.
\end{comment}
Between the start of openSNP on 09/27/2011 and 12/18/2011, 214 people have signed 
up with openSNP, 79 of which have uploaded their genotyping files. The openSNP 
database thereby lists 69,486,471 SNPs, which are distributed over 1,938,604 unique Rs-IDs.

\subsubsection*{Crowdsourcing phenotypes}
Beside uploading genotypic data, users may list new phenotypes that are not yet 
queried by openSNP. The specification of these phenotypes  is open and not limited 
to pre-defined categories. To reduce the amount of manual data curation, openSNP tries to harmonize 
the expression and spelling of the same phenotype or variation. We implemented an 
autocompletion-feature which helps users to reuse already entered phenotypes.
Users are encouraged to list as many phenotypes as possible through a simple 
achievement system which rewards users that upload their data and enter phenotypic 
information with small badges that are shown on their profile pages.
\begin{comment}
FZ: I changed the structure of the paragraph to make it a little bit more
concise.
I think its best to connect the phenotypes to the above paragraph, explain
the procedure and then talk about the achievments
\end{comment}
In the same timeframe that was mentioned above, all users combined have 
entered a total of 675 variations on 47 different phenotypes. See figure n for a 
distribution of data acquisition over time.
The mean number of users that have entered their variations for a single phenotype 
is 14.36 (SD 12.65), the median is 10. The distribution of how many users have 
entered their data per phenotype can be seen in figure n+1. The phenotype provided 
y the most users is the eye color, which has been entered by 54 users. There are 
two phenotypes which have so far only been provided by a single user: 
The SAT Writing score and triglyceride-levels.
\begin{comment}
FZ: I think we should clarify the difference between variaton and phenotype!
\end{comment}

>>>>

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW
\long\def\authornote#1{%
        \leavevmode\unskip\raisebox{-3.5pt}{\rlap{$\scriptstyle\diamond$}}%
        \marginpar{\raggedright\hbadness=10000
        \def\baselinestretch{0.8}\tiny
        \it #1\par}}
\newcommand{\bastian}[1]{\authornote{BG: #1}}
\newcommand{\fabian}[1]{\authornote{FZ: #1}}

@@

The openSNP project (http://opensnp.org) is an open, crowdsourced online platform for both DTC customers and researchers. DTC customers are encouraged to share their genotyping results and connect them with phenotypes to enable GWAS for researchers or any other reuse in a scientific context. 
\fabian{That is a good summary, but as Bastian pointed out the reuse is not only
enabled in a scientific context?}
\subsubsection*{Sharing genotypic information}
Currently users can upload their genotypings from the companies 23andme, deCODEme and Family Tree to/via a webinterface to the openSNP 
project. The uploaded data is published under the Creative Commons Zero-license, 
which - in accordance with the Panton Principles (doi:10.1371/journal.pbio.1001195) - 
allows a complete reuse of the data without any constraints.
\fabian{If that is true, rephrase sentence above. Made two sentences out of one.}
Between the start of openSNP on 09/27/2011 and 12/18/2011, 214 people have signed 
up with openSNP, 79 of which have uploaded their genotyping files. The openSNP 
database thereby lists 69,486,471 SNPs, which are distributed over 1,938,604 unique Rs-IDs.

\subsubsection*{Crowdsourcing phenotypes}
Beside uploading genotypic data, users may list new phenotypes that are not yet 
queried by openSNP. The specification of these phenotypes  is open and not limited 
to pre-defined categories. To reduce the amount of manual data curation, openSNP tries to harmonize 
the expression and spelling of the same phenotype or variation. We implemented an 
autocompletion-feature which helps users to reuse already entered phenotypes.
Users are encouraged to list as many phenotypes as possible through a simple 
achievement system which rewards users that upload their data and enter phenotypic 
information with small badges that are shown on their profile pages.
\fabian{
I changed the structure of the paragraph to make it a little bit more
concise.I think its best to connect the phenotypes to the above paragraph, explain
the procedure and then talk about the achievments.}

In the same timeframe that was mentioned above, all users combined have 
entered a total of 675 variations on 47 different phenotypes. See figure n for a 
distribution of data acquisition over time.
The mean number of users that have entered their variations for a single phenotype 
is 14.36 (SD 12.65), the median is 10. The distribution of how many users have 
entered their data per phenotype can be seen in figure n+1. The phenotype provided 
y the most users is the eye color, which has been entered by 54 users. There are 
two phenotypes which have so far only been provided by a single user: 
The SAT Writing score and triglyceride-levels.
\fabian{I think we should clarify the difference between variaton and phenotype!}
